NWOS Estimation, Revised Once Again
========================================================

During the week of Monday 18-Mar-13, we decided that the best way to handle both FIA and NWOS 
non-response might be to "calibrate" all estimates to FIA's estimates of area in the population of interest. I discovered, with unbridled dismay, that this is algebraically equivalent to the "old" estimators as presented (albeit incorrectly) in Butler et al 2005. See Dropbox/NWOS/brents nwos stuff/Documentation/EstimationCode-2013-03-25.docx for details.

```{r functions}
functions = list("MakeQuestAll"
                 ,"MakeAreaFia"
                 ,"StateTables"
                 ,"HheCore"
                 ,"CovCore"
                 ,"VarianceRatio"
                 ,"TweakQuest"
                 ,"RegionTables"
                 ,"StateTableContinuous"
                 ,"CalcSffi")
functions = sapply(functions, function(x) paste("functions/", x, ".R", sep = ""))
for(i in 1:length(functions)){source(functions[i])
                              rm(i)}
```

```{r questAll}
TweakQuest()
debugonce(MakeQuestAll)
QuestAll_11 = MakeQuestAll(quest = quest_11, NwosCycle = 2011)
QuestAll_06 = MakeQuestAll(quest = quest_06, NwosCycle = 2006)
AreaFia_11 = MakeAreaFia()[[1]]
AreaFia_06 = MakeAreaFia()[[2]]
```

```{r}
t23State_11 = StateTables(VariableColumns = which(grepl("allprivate", names(QuestAll_11)) & grepl("continuous", names(QuestAll_11)) == F)
                          ,Quest = QuestAll_11
                          ,PopulationOfInterest = "AllPrivate"
                          ,AreaFia = AreaFia_11
                          ,NwosCycle = 2011)
tContState_11 = StateTableContinuous(VariableColumns = which(grepl("continuous", names(QuestAll_11)) & grepl("allprivate", names(QuestAll_11)) == F)
                                      ,Quest = QuestAll_11
                                      ,DomainOfInterest = rep("domain", nrow(QuestAll_11))
                                      ,AreaFia = AreaFia_11
                                      ,NwosCycle = 2011)
t433State_11 = StateTables(which(grepl("__", names(QuestAll_11)) & grepl("continuous", names(QuestAll_11)) == F & grepl("allprivate", names(QuestAll_11)) == F)
                          ,Quest = QuestAll_11
                          ,PopulationOfInterest = "Family"
                          ,AreaFia = AreaFia_11
                          ,NwosCycle = 2011)
t433State_11 = rbind(t433State_11, tContState_11)
t23Region_11 = RegionTables(t23State_11)
t433Region_11 = RegionTables(t433State_11)
wd = "c:/users/ffrc_brent/dropbox/nwos/brents nwos stuff/output data/nwos tables/"
write.csv(t23Region_11, paste(wd, "Region_2011_AllPrivate_", substr(Sys.time(), 1, 10), ".csv", sep = ""), row.names = F)
write.csv(t23State_11, paste(wd, "State_2011_AllPrivate_", substr(Sys.time(), 1, 10), ".csv", sep = ""), row.names = F)
write.csv(t433Region_11, paste(wd, "Region_2011_Family_", substr(Sys.time(), 1, 10), ".csv", sep = ""), row.names = F)
write.csv(t433State_11, paste(wd, "State_2011_Family_", substr(Sys.time(), 1, 10), ".csv", sep = ""), row.names = F)
```


```{r}
t23State_06 = StateTables(VariableColumns = which(grepl("table_02", names(QuestAll_06)) | grepl("table_03", names(QuestAll_06)))
                          ,Quest = QuestAll_06
                          ,PopulationOfInterest = "AllPrivate"
                          ,AreaFia = AreaFia_06
                          ,NwosCycle = 2006)
tContState_06 = StateTableContinuous(VariableColumns = which(names(QuestAll_06) %in% c("table_NA__tenure_continuous"))
                                      ,Quest = QuestAll_06
                                      ,DomainOfInterest = rep("domain", nrow(QuestAll_06))
                                      ,AreaFia = AreaFia_06
                                      ,NwosCycle = 2006)
t433State_06 = StateTables(which(grepl("table_02", names(QuestAll_06)) == F & grepl("continuous", names(QuestAll_06)) == F & grepl("table_03", names(QuestAll_06)) == F & grepl("table_", names(QuestAll_06)) == T)
                          ,Quest = QuestAll_06
                          ,PopulationOfInterest = "Family"
                          ,AreaFia = AreaFia_06
                          ,NwosCycle = 2006)
t433State_06 = rbind(t433State_06, tContState_06)
```

Special tables for Marla:
```{r}
t433State_06 = StateTables(250
                          ,Quest = QuestAll_06
                          ,PopulationOfInterest = "Family"
                          ,AreaFia = AreaFia_06[AreaFia_06$StateCode == 1,]
                          ,NwosCycle = 2006)

```


```{r}
t23Region_06 = RegionTables(t23State_06)
t433Region_06 = RegionTables(t433State_06)

```

```{r}
wd = "c:/users/ffrc_brent/dropbox/nwos/brents nwos stuff/output data/nwos tables/"
write.csv(t23Region_06, paste(wd, "Region_2006_2-3_", substr(Sys.time(), 1, 10), ".csv", sep = ""), row.names = F)
write.csv(t23State_06, paste(wd, "State_2006_2-3_", substr(Sys.time(), 1, 10), ".csv", sep = ""), row.names = F)

write.csv(t433Region_06, paste(wd, "Region_2006_4-33_", substr(Sys.time(), 1, 10), ".csv", sep = ""), row.names = F)
write.csv(t433State_06, paste(wd, "State_2006_4-33_", substr(Sys.time(), 1, 10), ".csv", sep = ""), row.names = F)
```
